\documentclass[aspectratio=169]{beamer}  % 16:9 aspect ratio

% Use a clean theme as base
\usetheme{default}
\usecolortheme{default}

% Custom colors from HKUST logo
\definecolor{hkustblue}{RGB}{0, 51, 119}    % Navy blue from logo
\definecolor{hkustgold}{RGB}{180, 141, 61}  % Golden brown from logo
\definecolor{lightgray}{RGB}{236, 240, 241}

% Customize the appearance
\setbeamercolor{structure}{fg=hkustblue}
\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{normal text}{fg=hkustblue}
\setbeamercolor{frametitle}{fg=hkustblue,bg=white}
\setbeamercolor{itemize item}{fg=hkustgold}
\setbeamercolor{itemize subitem}{fg=hkustgold}
\setbeamercolor{block title}{fg=white,bg=hkustblue}
\setbeamercolor{block body}{fg=hkustblue,bg=lightgray}
\setbeamercolor{title}{fg=hkustblue}
\setbeamercolor{subtitle}{fg=hkustgold}

% Remove navigation symbols
\setbeamertemplate{navigation symbols}{}

% Customize frame title
\setbeamertemplate{frametitle}{
    \vspace*{0.5cm}
    \insertframetitle
    \vspace*{0.2cm}
    \begin{beamercolorbox}[wd=\paperwidth,ht=0.2pt]{structure}
    \end{beamercolorbox}
}

% Customize itemize bullets
\setbeamertemplate{itemize item}{\small\raise0.5pt\hbox{\textbullet}}
\setbeamertemplate{itemize subitem}{\tiny\raise1.5pt\hbox{\textbullet}}

% Packages
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
% add page number
\setbeamertemplate{footline}[frame number]

% Title page information
\title{Structural Analysis of Vertical Contracting}
\subtitle{Empirical Framework}
\author{Siyuan (Bruce) Jin,\\3rd Year in Information Systems}
\institute{Hong Kong University of Science and Technology}
\date{March, 2025}

\begin{document}

% Title page
\begin{frame}
    \titlepage
\end{frame}

% Slide 1: Introduction
\begin{frame}{Introduction}
    \begin{itemize}
        % \item This section focuses on an empirical framework for analyzing vertical contracting.
        \item Focus: theoretical insights into industry models to estimate firm behavior.
        \item Framework involves:
        \begin{enumerate}
            \item Supply-side actions (Stage 1)
            \item Demand-side actions (Stage 2)
        \end{enumerate}
        \item Outcomes modeled over single or multiple time periods.
    \end{itemize}
\end{frame}

% Slide 2: Two-Stage Framework
\begin{frame}{Two-Stage Framework for Analysis}
    \begin{itemize}
        \item Researchers often model vertical contracting using a two-stage framework.
        \item Stage 1: Supply
        \begin{itemize}
            \item Firms negotiate contracts and take payoff-relevant actions (e.g., investment, pricing).
        \end{itemize}
        \item Stage 2: Demand
        \begin{itemize}
            \item Consumers purchase products/services provided by upstream and downstream firms.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 5: Contracts
\begin{frame}{Contracts}
    \begin{itemize}
        \item Contracts represent agreements between upstream and downstream firms.
        \item Denoted as:
        \[
        C_{ij} \in \mathbb{C}
        \]
        \item Where:
        \begin{itemize}
            \item \(C_{ij}\): Contract between upstream firm \(i\) and downstream firm \(j\).
            \item \(\mathbb{C}\): Set of feasible contracts.
        \end{itemize}
        \item Null contract (\(C_0\)) represents the disagreement (no contract) outcome.
    \end{itemize}
\end{frame}

% Slide 6: Payoff-Relevant Actions
\begin{frame}{Payoff-Relevant Actions}
    \begin{itemize}
        \item Actions not explicitly specified in contracts but affect payoffs.
        \item Denoted by:
        \[
        a = \{a_0, a_1(\mathbb{C}, a_0)\}
        \]
        \item Where:
        \begin{itemize}
            \item \(a_0\): Actions chosen simultaneously with contracts.
            \item \(a_1(\mathbb{C}, a_0)\): Actions chosen after contracts and initial actions.
        \end{itemize}
        \item Examples:
        \begin{itemize}
            \item Downstream pricing, product availability.
            \item Effort provision or investment.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 7: Payoff Representation
\begin{frame}{Payoff Representation}
    \textbf{Detailed Representation:}
    \begin{itemize}
        \item Payoffs for each upstream firm \(i\) and downstream firm \(j\) are represented as:
        \[
        \Pi_{U_i}(\mathbb{C}, \boldsymbol{a}_0), \quad \Pi_{D_j}(\mathbb{C}, \boldsymbol{a}_0)
        \]
        \item These payoffs implicitly depend on:
        \begin{itemize}
            \item Subsequent actions taken by firms (\(\boldsymbol{a}_1(\cdot)\)).
            \item Consumer actions captured by demand functions.
        \end{itemize}
    \end{itemize}

    \vspace{0.3cm}
    \textbf{Demand Representation:}
    \begin{itemize}
        \item Upstream demand: \(\bar{D}(\cdot)\).
        \item Downstream demand: \(\underline{D}(\cdot)\).
    \end{itemize}
\end{frame}


\begin{frame}{Payoff Representation}
  
    \textbf{Demand Representation:}
    \begin{itemize}
        \item Example (successive monopoly with per-unit pricing):
        \[
        \mathbb{C} = \{w\}, \quad a_0 = \emptyset, \quad a_1 = \{p\}
        \]
        \[
        \bar{D}(w) = \underline{D}(w) = D\left(p^m\left(w + c_R\right)\right)
        \]
        \item In this case, upstream and downstream demand coincide because of a single retailer-manufacturer setup.
    \end{itemize}

    \vspace{0.3cm}
    \textbf{General Case:}
    \begin{itemize}
        \item When there are multiple upstream firms or consumers do not always purchase upstream products, \(\bar{D}(\cdot)\) and \(\underline{D}(\cdot)\) will typically differ.
    \end{itemize}
\end{frame}

% Slide 8: Example - Medical Devices
\begin{frame}{Example: Medical Devices (Grennan, 2013)}
    \begin{itemize}
        \item Contracts \(\mathbb{C}\) between stent manufacturers (\(i\)) and hospitals (\(j\)) specify linear prices \(\boldsymbol{w} = \{w_{ij}\}\).
        \item The null contract \(\mathbb{C}_0\) is represented by \(w_{ij} = \infty\), meaning no trade occurs.
    \end{itemize}
\end{frame}

\begin{frame}{Example: Medical Devices (Grennan, 2013)}
    \begin{itemize}
        
        \item The payoff for stent manufacturer \(i\) is:
        \[
        \Pi_{U_i}(\mathbb{C}) = \sum_{j : \mathbb{C}_{ij} \neq \mathbb{C}_0} \left(w_{ij} - c_i\right) \bar{D}_{ij}\left(\{w_{kj}\}_{k \in \mathcal{I}}\right)
        \]
        \begin{itemize}
            \item \(w_{ij}\) is the price in the contract between manufacturer \(i\) and hospital \(j\).
            \item \(c_i\) is the marginal cost of producing stent \(i\).
            \item \(\bar{D}_{ij}(\cdot)\) is the quantity of stent \(i\) used at hospital \(j\), which depends on:
            \begin{itemize}
                \item Preferences of doctors and patients at hospital \(j\).
                \item Contracts \(\{w_{kj}\}\) for all stents signed by hospital \(j\).
            \end{itemize}
        \end{itemize}

        \item The payoff for hospital \(j\) is:
        \[
        \Pi_{D_j}(\mathbb{C}) = W_j\left(\{w_{kj}\}_{k \in \mathcal{I}}\right)
        \]
        \begin{itemize}
            \item \(W_j(\cdot)\) is the welfare of hospital \(j\), which depends on the prices \(\{w_{kj}\}\) for all stents it uses.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 1: Commercial Health Insurers and Hospitals - Setup
\begin{frame}{Example:  Health Insurers and Hospitals (Ho and Lee, 2017)}
    \begin{itemize}
        \item Contracts between hospitals (upstream) and insurers (downstream) in the U.S. healthcare industry.
        \item Contracts \(\mathbb{C}\) specify payments per hospital admission:
        \[
        \boldsymbol{w} = \{w_{ij}\}
        \]
        \item Insurers also set premiums for households:
        \[
        \boldsymbol{p} = \{p_j\}
        \]
        \item Demand terms:
        \begin{itemize}
            \item \(\underline{D}_j(\boldsymbol{p}, \boldsymbol{N})\): Households enrolled in insurer \(j\).
            \item \(\bar{D}_{ij}(\boldsymbol{p}, \boldsymbol{N})\): Admissions from insurer \(j\)'s enrollees to hospital \(i\).
        \end{itemize}
        \item \(\boldsymbol{N} = \{ij : \mathbb{C}_{ij} \neq \mathbb{C}_0\}\) represents the network of contracts.
    \end{itemize}
\end{frame}

% Slide 2: Commercial Health Insurers and Hospitals - Payoffs
\begin{frame}{Example: Health Insurers and Hospitals (Ho and Lee, 2017)}
    \begin{itemize}
        \item Hospital \(i\)'s payoff:
        \[
        \Pi_{U_i}(\mathbb{C}, \boldsymbol{p}) = \sum_{j : \mathbb{C}_{ij} \neq \mathbb{C}_0} \left(w_{ij} - c_i\right) \bar{D}_{ij}(\boldsymbol{p}, \boldsymbol{N})
        \]
        \begin{itemize}
            \item \(w_{ij}\): Payment per admission.
            \item \(c_i\): Per-admission cost.
        \end{itemize}
        
        \item Insurer \(j\)'s payoff:
        \[
        \Pi_{D_j}(\mathbb{C}, \boldsymbol{p}) = \left(p_j - \eta_j\right) \underline{D}_j(\boldsymbol{p}, \boldsymbol{N}) - \sum_{i : \mathbb{C}_{ij} \neq \mathbb{C}_0} w_{ij} \bar{D}_{ij}(\boldsymbol{p}, \boldsymbol{N})
        \]
        \begin{itemize}
            \item \(p_j\): Insurer premium.
            \item \(\eta_j\): Non-hospital costs (e.g., physician or drug payments).
        \end{itemize}
        
        \item Differences from the previous example:
        \begin{itemize}
            \item Payoffs depend on premiums, an additional supply-side decision.
            \item Insurers compete for households, so all firms’ actions affect payoffs.
        \end{itemize}
    \end{itemize}
\end{frame}


% Slide 1: Multichannel Television - Setup
\begin{frame}{Example: Multichannel Television and Vertical Integration}
    \begin{itemize}
        \item Crawford and Yurukoglu (2012) and Crawford et al. (2018) study negotiations between:
        \begin{itemize}
            \item Upstream television channels (\(i\)).
            \item Downstream multichannel video programming distributors (MVPDs, \(j\)), such as cable and satellite firms.
        \end{itemize}
        \item Contracts \(\mathbb{C}\) specify linear affiliate fees:
        \[
        \boldsymbol{w} = \{w_{ij}\}
        \]
        \item \(w_{ij}\): Amount distributor \(j\) pays channel \(i\) per subscriber.
        \item Distributors choose subscription prices:
        \[
        \boldsymbol{p} = \{p_j\}
        \]
    \end{itemize}
\end{frame}

% Slide 2: Multichannel Television - Timing and Demand
\begin{frame}{Timing and Demand in Multichannel Television}
    \begin{itemize}
        \item Timing assumptions:
        \begin{itemize}
            \item Crawford and Yurukoglu (2012): Subscription prices are set after contracting.
            \item Crawford et al. (2018): Subscription prices and contracts are set simultaneously.
        \end{itemize}
        \item Demand objects:
        \begin{itemize}
            \item \(\underline{D}_j(\boldsymbol{p}, \boldsymbol{N})\): Number of households subscribing to distributor \(j\).
            \item \(\boldsymbol{N} = \{ij : \mathbb{C}_{ij} \neq \mathbb{C}_0\}\): Network of channel-distributor agreements.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 3: Multichannel Television - Payoffs
\begin{frame}{Payoffs in Multichannel Television}
    \begin{itemize}
        \item Channel \(i\)'s payoff:
        \[
        \Pi_{U_i}(\mathbb{C}, \boldsymbol{p}) = \sum_{j : \mathbb{C}_{ij} \neq \mathbb{C}_0} \left(w_{ij} \underline{D}_j(\boldsymbol{p}, \boldsymbol{N}) + ad_{ij}(\boldsymbol{p}, \boldsymbol{N})\right)
        \]
        \begin{itemize}
            \item \(ad_{ij}(\cdot)\): Advertising revenue from distributor \(j\)'s subscribers.
            \item $w_{ij}$: channel fee received from distributor $j$.
        \end{itemize}

        \item Distributor \(j\)'s payoff:
        \[
        \Pi_{D_j}(\mathbb{C}, \boldsymbol{p}) = \left(p_j - \sum_{i : \mathbb{C}_{ij} \neq \mathbb{C}_0} w_{ij}\right) \underline{D}_j(\boldsymbol{p}, \boldsymbol{N})
        \]
        \begin{itemize}
            \item \(p_j\): Subscription price set by distributor \(j\).
            \item \(\underline{D}_j(\cdot)\): Number of households subscribing to distributor \(j\).
        \end{itemize}

    \end{itemize}
\end{frame}

\begin{frame}{Discussion}
    \begin{itemize}
            \item Key difference from Example 11:
        \begin{itemize}
            \item Here, fees \(w_{ij}\) are paid for all subscribers (\(\underline{D}_j\)).
            \item In Example 11, fees were paid only for specific hospital admissions (\(\bar{D}_{ij}\)).
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 13: Sequential vs Simultaneous Timing
\begin{frame}{Sequential vs Simultaneous Timing}
    \begin{itemize}
        \item Timing assumptions play a critical role in contracting models:
        \begin{itemize}
            \item Simultaneous: Actions like pricing and contracting are decided together.
            \item Sequential: Contracting concludes before other actions are taken (e.g., pricing).
        \end{itemize}
        \item Implications:
        \begin{itemize}
            \item Simultaneous timing simplifies calculations but ignores downstream adjustments.
            \item Sequential timing captures adjustment dynamics but increases computational complexity.
        \end{itemize}
        \item Example: Multichannel TV contracts often assume sequential timing for pricing decisions.
    \end{itemize}
\end{frame}

% Slide 14: Modeling Contract Formation
\begin{frame}{Modeling Contract Formation}
    \begin{itemize}
        \item Different approaches to modeling contract formation:
        \begin{enumerate}
            \item Take-It-Or-Leave-It (TIOLI) Offers:
            \begin{itemize}
                \item One side (e.g., upstream firms) makes offers to the other side.
                \item Example: Villas-Boas (2007) for manufacturer-retailer contracts.
            \end{itemize}
            \item Nash-in-Nash Bargaining:
            \begin{itemize}
                \item Firms bargain simultaneously over terms.
                \item Useful for capturing mutual leverage and flexibility in negotiations.
            \end{itemize}
        \end{enumerate}
        \item Choice of model depends on the industry and available data.
    \end{itemize}
\end{frame}

% Slide 1: Contracting Between Manufacturers and Retailers - Setup
\begin{frame}{Contracting Between Manufacturers and Retailers}
    \begin{itemize}
        \item Villas-Boas (2007) models contracting and pricing between:
        \begin{itemize}
            \item Manufacturers (yogurt producers, indexed by \(i\)).
            \item Retailers (supermarkets, indexed by \(j\)).
        \end{itemize}
        \item Two stages:
        \begin{itemize}
            \item Stage 1: Manufacturers simultaneously offer contracts to retailers.
            \item Stage 2: Retailers choose retail prices given contracts.
        \end{itemize}
        \item Contracts may include:
        \begin{itemize}
            \item Linear wholesale prices \(\boldsymbol{w} = \{w_{ij}\}\).
            \item Two-part tariffs (fixed fees and variable prices).
        \end{itemize}
        \item Assumptions:
        \begin{itemize}
            \item Each manufacturer produces one product.
            \item Each retailer carries all products.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 2: Retailer Profit and Pricing
\begin{frame}{Retailer Profit and Pricing}
    \begin{itemize}
        \item Retailer \(j\)'s profit:
        \[
        \Pi_{D_j}(\boldsymbol{w}, \boldsymbol{p}) = \sum_i \left(p_{ij} - w_{ij} - c_{Dj}\right) D_{ij}(\boldsymbol{p})
        \]
        \begin{itemize}
            \item \(p_{ij}\): Retail price of product \(i\) at store \(j\).
            \item \(w_{ij}\): Wholesale price offered by manufacturer \(i\).
            \item \(c_{Dj}\): Retailer \(j\)'s marginal cost.
            \item \(D_{ij}(\boldsymbol{p})\): Demand for product \(i\) at store \(j\).
        \end{itemize}
        \item Retail prices under Nash-Bertrand competition:
        \[
        \boldsymbol{p} = \boldsymbol{w} + \boldsymbol{c}_D - \left(\boldsymbol{T} * \Lambda_D\right)^{-1} \boldsymbol{D}(\boldsymbol{p})
        \]
        \begin{itemize}
            \item \(\boldsymbol{T}(m, n)\): 1 if \(m, n\) share the same retailer, 0 otherwise.
            \item \(\Lambda_D(m, n) = \frac{\partial D_n}{\partial p_m}\): Demand sensitivity to price changes.
            \item \( * \): Element-wise multiplication.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 3: Manufacturer Profit
\begin{frame}{Manufacturer Profit}
    \begin{itemize}
        \item Manufacturer \(i\)'s profit:
        \[
        \Pi_{U_i}(\boldsymbol{w}) = \sum_j \left(w_{ij} - c_{Ui}\right) D_{ij}(\boldsymbol{p}(\boldsymbol{w}))
        \]
        \begin{itemize}
            \item \(w_{ij}\): Wholesale price offered to retailer \(j\).
            \item \(c_{Ui}\): Manufacturer \(i\)'s marginal cost.
            \item \(D_{ij}(\boldsymbol{p}(\boldsymbol{w}))\): Demand for product \(i\) at store \(j\), given equilibrium prices.
        \end{itemize}
        \item Equilibrium wholesale prices:
        \[
        \boldsymbol{w} = \boldsymbol{c}_U - \left(I_{I \times J} * (\Lambda_P' \Lambda_D)\right)^{-1} \boldsymbol{D}(\boldsymbol{p}(\boldsymbol{w}))
        \]
        \begin{itemize}
            \item \(I_{I \times J}\): Identity matrix for all product-retailer combinations.
            \item \(\Lambda_P(m, n) = \frac{\partial p_n}{\partial w_m}\): Pass-through matrix.
        \end{itemize}
        \item Villas-Boas uses this framework to analyze price-cost margins and conduct assumptions.
    \end{itemize}
\end{frame}

% Slide 16: Nash-in-Nash Bargaining Model
\begin{frame}{Nash-in-Nash Bargaining Model}
    \begin{itemize}
        \item Nash-in-Nash bargaining captures simultaneous negotiations between pairs of firms.
        \item Necessary condition:
\begin{footnotesize}
\[
\widehat{\mathbb{C}}_{ij} \in \arg \max_{\mathbb{C}_{ij} \in \mathcal{C}_{ij}^+(\widehat{\mathbb{C}}_{-ij})}  
\Bigg[
\underbrace{
    \Big(\Pi_{Dj}(\{\mathbb{C}_{ij}, \widehat{\mathbb{C}}_{-ij}\}) 
    - \Pi_{Dj}(\{\mathbb{C}_0, \widehat{\mathbb{C}}_{-ij}\})\Big)^{b_{ij}}
}_{\Delta_{ij} \Pi_{Dj}(\{\mathbb{C}_{ij}, \widehat{\mathbb{C}}_{-ij}\})}
\cdot
\underbrace{
    \Big(\Pi_{Ui}(\{\mathbb{C}_{ij}, \widehat{\mathbb{C}}_{-ij}\}) 
    - \Pi_{Ui}(\{\mathbb{C}_0, \widehat{\mathbb{C}}_{-ij}\})\Big)^{1-b_{ij}}
}_{\Delta_{ij} \Pi_{Ui}(\{\mathbb{C}_{ij}, \widehat{\mathbb{C}}_{-ij}\})}
\Bigg]
\]
\end{footnotesize}
        \item Key terms:
        \begin{itemize}
            \item \(\Delta D_j(C)\): Gains from trade for downstream firm \(j\).
            \item \(\Delta U_i(C)\): Gains from trade for upstream firm \(i\).
            \item \(b_{ij}\): Bargaining parameter for downstream firm \(j\).
        \end{itemize}
        \item Assumes contracts of other pairs are held fixed during negotiations.
    \end{itemize}
\end{frame}
% Slide 1: Health Insurer-Hospital Negotiations - Setup
\begin{frame}{Health Insurer-Hospital Negotiations (Ho and Lee, 2017)}
    \begin{itemize}
        \item Health insurers negotiate contracts \(\mathbb{C}\) with hospitals, specifying per-admission payments \(w\), while simultaneously negotiating premiums \(\boldsymbol{p}\) with employers.
        \item Payoffs are based on Example 11 (\(\Pi_{U_i}\) for hospitals and \(\Pi_{D_j}\) for insurers).
    \end{itemize}
\end{frame}


% Slide 2: Commercial Health Insurers and Hospitals - Payoffs
\begin{frame}{Example: Health Insurers and Hospitals (Ho and Lee, 2017)}
    \begin{itemize}
        \item Hospital \(i\)'s payoff:
        \[
        \Pi_{U_i}(\mathbb{C}, \boldsymbol{p}) = \sum_{j : \mathbb{C}_{ij} \neq \mathbb{C}_0} \left(w_{ij} - c_i\right) \bar{D}_{ij}(\boldsymbol{p}, \boldsymbol{N})
        \]
        \begin{itemize}
            \item \(w_{ij}\): Payment per admission.
            \item \(c_i\): Per-admission cost.
        \end{itemize}
        
        \item Insurer \(j\)'s payoff:
        \[
        \Pi_{D_j}(\mathbb{C}, \boldsymbol{p}) = \left(p_j - \eta_j\right) \underline{D}_j(\boldsymbol{p}, \boldsymbol{N}) - \sum_{i : \mathbb{C}_{ij} \neq \mathbb{C}_0} w_{ij} \bar{D}_{ij}(\boldsymbol{p}, \boldsymbol{N})
        \]
        \begin{itemize}
            \item \(p_j\): Insurer premium.
            \item \(\eta_j\): Non-hospital costs (e.g., physician or drug payments).
        \end{itemize}
        
        \item Differences from the previous example:
        \begin{itemize}
            \item Payoffs depend on premiums, an additional supply-side decision.
            \item Insurers compete for households, so all firms’ actions affect payoffs.
        \end{itemize}
    \end{itemize}
\end{frame}


% Slide 1: Health Insurer-Hospital Negotiations - Setup
\begin{frame}{Health Insurer-Hospital Negotiations (Ho and Lee, 2017)}
    \begin{itemize}
        \item Nash-in-Nash bargaining conditions govern hospital payments \(w_{ij}\), incorporating premiums \(\boldsymbol{p}\) as supply-side actions (first-order condition):
        \[
        \widehat{w}_{ij} \times \bar{D}_{ij}(\cdot) = 
        \left(1 - b_{ij}\right) \times \Delta_{ij} \Pi_{D_j}\left(\{w_{ij} = 0, \widehat{\mathbb{C}}_{-ij}\}, \widehat{\boldsymbol{p}}\right)
        - b_{ij} \times \Delta_{ij} \Pi_{U_i}\left(\{w_{ij} = 0, \widehat{\mathbb{C}}_{-ij}\}, \widehat{\boldsymbol{p}}\right)
        \]
        \item Terms in the equation:
        \begin{itemize}
            \item \(\bar{D}_{ij}(\cdot)\): Number of insurer \(j\)'s enrollees admitted to hospital \(i\).
            \item \(b_{ij}\): Bargaining weight of hospital \(i\) in negotiations with insurer \(j\).
            \item \(\Delta_{ij} \Pi_{D_j}\): Insurer \(j\)'s gain from trade if the payment to hospital \(i\) is set to zero.
            \item \(\Delta_{ij} \Pi_{U_i}\): Hospital \(i\)'s gain from trade under the same condition.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 18: Double Marginalization
% \begin{frame}{Double Marginalization}
%     \begin{itemize}
%         \item Double marginalization occurs when both upstream and downstream firms add markups.
%         \item Example: Manufacturer-retailer chain.
%         \item Nash bargaining reduces double marginalization by sharing surplus.
%         \item Equilibrium wholesale price:
%         \[
%         w = (1 - b) p + b c
%         \]
%         \begin{itemize}
%             \item \(b\): Bargaining parameter for downstream firm.
%             \item \(p\): Retail price.
%             \item \(c\): Upstream marginal cost.
%         \end{itemize}
%         \item Retail price still exceeds marginal cost, but markups are lower.
%     \end{itemize}
% \end{frame}

% Slide 19: Estimation and Identification
\begin{frame}{Supply Estimation and Identification}
    \begin{itemize}
        \item Estimation involves recovering key parameters:
        \begin{itemize}
            \item Marginal costs (\(c^U, c^D\)).
            \item Bargaining parameters (\(b_{ij}\)).
            \item Gains from trade (\(\Delta D_j, \Delta U_i\)).
        \end{itemize}
        \item Ideal data scenario:
        \begin{itemize}
            \item Observed wholesale prices (\(w\)).
            \item Observed demand system (\(D(p)\)).
            \item Marginal cost $c_R$ and $c_M$.
        \end{itemize}
        \item Missing data (e.g., marginal costs) requires additional assumptions or instruments.
    \end{itemize}
\end{frame}

% Slide 21: Demand Estimation
\begin{frame}{Demand Estimation}
    \begin{itemize}
        \item A key input for vertical contracting models is estimating consumer demand.
        \item Demand estimation helps predict:
        \begin{itemize}
            \item Upstream and downstream quantities (\(D(p, x, w)\)).
            \item Consumer responses to prices and product characteristics (\(x\)).
        \end{itemize}
        \item Techniques:
        \begin{itemize}
            \item Use exogenous variation in prices and characteristics to identify demand.
            \item Estimate demand functions for both upstream and downstream firms.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 22: Example - Estimating Demand in Healthcare
\begin{frame}{Example: Estimating Demand in Healthcare}
    \begin{itemize}
        \item Ho and Lee (2017): Model demand for health insurance plans and hospital services.
        \item Consumer utility for insurer \(j\):
        \[
        u_{cjm} = \beta v_{cjm} + x_{jm} \beta_x + \xi_{jm} + \epsilon_{cjm}
        \]
        \begin{itemize}
            \item \(v_{cjm}\): Willingness to pay (WTP) for insurer \(j\)'s hospital network.
            \item \(x_{jm}\): Observed characteristics (e.g., premiums).
            \item \(\xi_{jm}\): Unobserved demand shocks.
            \item \(\epsilon_{cjm}\): Idiosyncratic preferences.
        \end{itemize}
        \item Model jointly estimates insurer and hospital demand.
    \end{itemize}
\end{frame}

% Slide 23: Usage Models for Bundles
\begin{frame}{Usage Models for Bundles}
    \begin{itemize}
        \item When consumers purchase bundles, usage data can inform valuation.
        \item Example: Multichannel TV (Crawford et al., 2018)
        \begin{itemize}
            \item Consumer utility for distributor \(j\):
            \[
            u_{cjm} = \beta v_{cjm}(C_j) + x_{jm} \beta_x + \xi_{jm} + \epsilon_{cjm}
            \]
            \item \(v_{cjm}(C_j)\): Viewership utility for channels in bundle \(C_j\).
            \item Viewership data helps estimate valuations for individual channels.
        \end{itemize}
        \item Usage models reduce data requirements by linking upstream and downstream choices.
    \end{itemize}
\end{frame}

% Slide 24: Upstream Choice-Only Models
\begin{frame}{Upstream Choice-Only Models}
    \begin{itemize}
        \item In some cases, only upstream demand is modeled.
        \item Example: Grennan (2013) - Medical devices
        \begin{itemize}
            \item Focuses on hospitals' choice of medical devices.
            \item Does not model patient flows across hospitals.
        \end{itemize}
        \item Simplifies computation but ignores some competitive effects.
    \end{itemize}
\end{frame}

% Slide 25: Consumer Selection
\begin{frame}{Consumer Selection in Demand Estimation}
    \begin{itemize}
        \item Selection bias arises when observed consumption depends on unobserved preferences.
        \item Example: Multichannel TV (Crawford and Yurukoglu, 2012)
        \begin{itemize}
            \item Consumers who purchase bundles may have higher valuations for included channels.
            \item Ignoring selection leads to overestimating valuations.
        \end{itemize}
        \item Solution:
        \begin{itemize}
            \item Jointly estimate upstream and downstream demand.
            \item Use techniques like Heckman correction or joint estimation of preferences.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 26: Joint Estimation of Demand and Supply
\begin{frame}{Joint Estimation of Demand and Supply}
    \begin{itemize}
        \item Demand and supply parameters can be jointly estimated for efficiency.
        \item Example: Crawford et al. (2018)
        \begin{itemize}
            \item Wholesale prices (\(w_{ij}\)) used to infer demand valuations.
            \item Assumes content with higher fees has higher consumer value.
        \end{itemize}
        \item Benefits:
        \begin{itemize}
            \item Increases precision of demand estimates.
            \item Captures interactions between demand and supply.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 27: Policy Applications - Mergers
\begin{frame}{Policy Applications: Horizontal Mergers in Vertical Markets}
    \begin{itemize}
        \item Structural models are used to analyze merger impacts.
        \item Example: Gowrisankaran et al. (2015)
        \begin{itemize}
            \item Study hospital mergers and their effects on insurer-hospital bargaining.
            \item Simulate price increases under Nash-in-Nash bargaining.
        \end{itemize}
        \item Policy implications:
        \begin{itemize}
            \item Mergers can increase bargaining leverage but harm consumer welfare.
            \item Quantifying competitive effects is crucial for antitrust decisions.
        \end{itemize}
    \end{itemize}
\end{frame}

% Slide 28: Policy Applications - Vertical Integration
\begin{frame}{Policy Applications: Vertical Integration}
    \begin{itemize}
        \item Vertical integration affects competition through:
        \begin{itemize}
            \item Efficiencies: Reducing double marginalization.
            \item Foreclosure: Denying rivals access to key inputs.
        \end{itemize}
        \item Example: Crawford et al. (2018)
        \begin{itemize}
            \item Study integration between cable distributors and sports channels.
            \item Find positive average welfare effects but harm to rival distributors.
        \end{itemize}
        \item Policy implications:
        \begin{itemize}
            \item Effects of vertical integration vary across markets.
            \item Requires case-specific analysis for mergers.
        \end{itemize}
    \end{itemize}
\end{frame}



\end{document} 